Model-agnostic Interpretable Machine Learning

### 2021-08-21 2021-bips-practice-phd-defence/2021-bips-practice-phd-defence.Rmd
continues draft

- A Brief History of Interpretable Modeling
- Model-agnostic interpretability
- Methods PDP and PFI
- What I did in my PhD
    - Conditional Feature for PFI and PDP Importance Using Subgroups
    - Statistical Inference for PDP and PFI
    - Measuring Model Complexity

## Before You Pin Me Down

To get this out of the way, here is how I use "Interpretable Machine Learning":

It's an umbrella term for approaches that analyze or describe prediction mechanisms in models that were learned on data.

@@

* 1800s: Linear regression models
* 1960s: Rule-based ML
* 2001:  Random Forest
* 2012:  Deep Learning Boom kicked off
* 2016:  Interpretable ML / Explainable AI boom

@@

TODO: Add citation report and or google trends image

@@

* The more non-linear and complex a model, the more difficult to understand how it operates
* Trust
* Debugging
* Control
* ...

@@

* GAM and other linear models
* Trees and decision rules
* More complex models, but interpretation requires more work or post-processing

@@

@@

## Interpreting Models

@@

## Permutation Feature Importance PFI

The construct:
$$PFI =  \mathbb{E}_{Y \tilde{X}_S X_C}(L(Y, \hat{f}(\tilde{X}_S, X_C))) - \mathbb{E}_{Y X_S X_C}(L(Y,\hat{f}(X_S, X_C)))$$

@@

## PFI and PDP similarity

CITE paper GC here

@@

# Summary of Thesis

TODO: Draw image with Ipad

In my thesis I did a lot of consolidation work:

- Book
- R package
- History paper
- Pitfalls paper

What if we embrace black box view of models?

Special focus on old, established tools, and make them fit for right use in practice

Also some more specific papers

* subgroups
* inference
* quantifying

and some contributions to

* MOC
* relative feature importance
* SIPA
* feature importance and PDP (GC)

>>>>

- Introduction
  - Brief History of Interpretability
  - What Is Model-Agnostic Interpretability?
  - Partial Dependence Plot (PDP)
  - Permutation Feature Importance (PFI)
- What I did in my PhD
- Statistical Inference for PDP and PFI

# Introduction

@@

@@

>- 1800s: Linear regression models
>- 1960s: Rule-based ML
>- 2001:  Random Forest
>- 2012:  Deep Learning Boom kicked off
>- 2016:  Interpretable ML / Explainable AI boom

TODO: Web Of Science for keywords XAI and IML (written out)

@@

```{r, out.width = 500}
knitr::include_graphics("../images/explain.png")
```

@@

* Decision trees and decision rules
* Generalized additive models and the whole gang of statistical modeling
* Interpretable models allow interpretation of internal structure and parameters.

@@

## Black Box Models

Random forest, gradient-boosted trees, neural networks, ...

@@

@@

## Permutation Feature Importance (PFI)

How important is a feature for making correct predictions?

## Permutation Feature Importance (PFI)

How important is a feature for making correct predictions?

Idea: Shuffle the feature, and see how much worse the predictions are.

```{r}
knitr::include_graphics("../images/permimp-shuffle.png")
```

## Permutation Feature Importance (PFI)

The construct:

$$PFI =  \mathbb{E}_{Y \tilde{X}_S X_C}(L(Y, \hat{f}(\tilde{X}_S, X_C))) - \mathbb{E}_{Y X_S X_C}(L(Y,\hat{f}(X_S, X_C)))$$

@@

@@

# Stuff I Did for My Thesis

## Most Important

## Not Part of My PhD Though

```{r, out.width = 540}
knitr::include_graphics("../images/academic-credit.jpg")
```

## Summary of Thesis

- Consolidation Work
  - Book
  - R package
  - History paper
  - Pitfalls paper
- Improving Methods
  - subgroups
  - iml inference
  - quantifying
- Contributed to
  - MOC
  - relative feature importance
  - SIPA
  -  feature importance and PDP (GC)
